---
title: "Project2-data3"
author: "Kossi Akplaka"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

## Import libraries
```{r import-libraries, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
```

## Collect the data
This data was collected on Slack from the student: FOMBA KASSOH.

```{r}
df3 = read.csv("https://raw.githubusercontent.com/Kossi-Akplaka/Data607-data_acquisition_and_management/main/Project%202/Data3-USA-Development-Indicators.csv")
kable(head(df3)) 
```
## Let's tidy the data
Remove columns 56 to 60
```{r}
df3 <- df3[1:55,]
```

We can transform the data from wide to long

```{r}
df3_long <- df3 %>%
  gather(Year, Count, X1990..YR1990.:X2022..YR2022.)
kable(head(df3_long))
```
Now we can tidy the column Year
```{r}
df3_long$Year <- as.integer(sub("X(\\d+)\\.\\.YR\\d+\\.", "\\1", df3_long$Year))

head(df3_long$Year)
```
## Analysis
Let's create another data frame with the total population of USA

```{r}
USA_pop_total <- df3_long %>% 
  filter(Series.Name == "Population, total")
kable(USA_pop_total)
```

Now we can plot the total population of USA between 1990 to 2022

```{r}
ggplot(data = USA_pop_total, aes(x = Year, y = Count, group = 1)) +
  geom_line()
```

The population of USA grew from `r min(USA_pop_total$Count)` to `r max(USA_pop_total$Count)`.


