---
title: "Project2-data2"
author: "Kossi Akplaka"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

## Import libraries
```{r import-libraries, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library('zoo') # fill the missing values with the previous non missing value
library(ggplot2)
```

## Collect the data
This data was posted on Slack by the student: Marley Myrianthopoulos
```{r}
df2 = read.csv("https://raw.githubusercontent.com/Kossi-Akplaka/Data607-data_acquisition_and_management/main/Project%202/Data2-MajorLeagueBaseballcsv.csv")
head(df2,7)
```
## Tidy the data
First we need to fill the missing values on the first column
```{r}
df2$Team[df2$Team == ""] <- NA
df2$Team <- na.locf(df2$Team)
df2
```
Now we can transform the data from wide to long 
```{r}
df2_long <- df2 %>% 
  pivot_longer(cols = c(X2018.HRs, X2019.HRs, X2021.HRs,
                        X2022.HRs,X2023.HRs) ,
               names_to = "Year",
               values_to = "Home Runs")

head(df2_long, 7)
```
Finally let's change the format for the year column
```{r}
df2_long$Year <- sub("^X", "", df2_long$Year)
df2_long$Year <- gsub(".HRs", "", df2_long$Year)
class(df2_long$Year)
```

Convert the year column to a date
```{r}
df2_long$Year <- as.integer(df2_long$Year)
head(df2_long,7)
```
## Analysis

Let's create a plot showing how many home runs were scored for the position "2B" by the team "BAL" and the team "NYY" overtime.
```{r}

df2_BAL <- df2_long %>% 
  filter(Team == 'BAL' & Position == '2B')
df2_NYY <- df2_long %>% 
  filter(Team == 'NYY' & Position == '2B')


ggplot() +
  geom_line(data = df2_BAL, aes(x = Year, y = `Home Runs`, color = "BAL"))+
  geom_line(data = df2_NYY, aes(x = Year, y = `Home Runs`, color = "NYY")) +
  labs(x = "Year", y = "Home Runs", title = "Home Runs Comparison")
 
```

The plot shows that NYY's players at Position 2B score more home runs than BAL's players.

