---
title: "Project 2- data1"
author: "Kossi Akplaka"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

## Import libraries
```{r import-libraries, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(qcc) # for pareto-analysis
```


## Analysis of DATA 1
I got the first data from the student: Sean Amato. I saved that on my github  after I downloaded that on Slack.

### Collect the data
```{r collect-data}
df1 = read.csv('https://raw.githubusercontent.com/Kossi-Akplaka/Data607-data_acquisition_and_management/main/Project%202/Data1-Manufacturing.csv', header = TRUE)
df1
```

### Tidy the data
First we can remove the header of the data frame and make the second column as a header
```{r remove-add-header}
colnames(df1) <- as.character(unlist(df1[1, ]))
df1 <- df1[-1, ] 
df1
```
Now, we can fill up the missing value for Building
```{r missing-values}
df1[2,1] = 'B7'
df1[4,1] = 'B23'
df1[6,1] = 'B30'

df1
```

Now, we can transform the data from wide to long format.

```{r wide-to-long}
colnames(df1)

df1_long <- df1 %>% 
  pivot_longer(cols = c( "NOASSIGN",  "WORKASSIGN","WORKPLAN",
                         "WORKSCHED"  ,"INPROG"  ,   "COMP"),
               names_to = "Order_Status",
               values_to = "Count")

head(df1_long, 7)
```

### Data analysis
We have a tidy data ready for analysis. We can plot a Pareto chart. 
you can find more detail how to create a Pareto chart in this [article]("https://www.statology.org/pareto-chart-in-r/")
```{r pareto-chart}
df1_long$Count <- as.numeric(df1_long$Count)
pareto.chart(df1_long$Count)
```



