---
title: "Assignment 3:  R Character Manipulation and Date Processing"
author: "Kossi Akplaka"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

Please deliver links to an R Markdown file (in GitHub and rpubs.com) with solutions to the problems below.  You may work in a small group, but please submit separately with names of all group participants in your submission.

## 1. Provide code that identifies the majors that contain either "DATA" or "STATISTICS"

Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/] 

```{r load-data}
majors_df = read.csv('https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/recent-grads.csv', header = TRUE)
```

Create a pattern that identifies the major containing either "DATA" or "STATISTICS"
```{r return-data-stats-major}
pattern <- "(DATA|STATISTICS)"
matches <- grep(pattern, majors_df$Major, value = TRUE, ignore.case = TRUE)
matches
```

## 2 Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")


```{r transformed-data}
raw_data <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"'

transformed_data <- scan(text = raw_data, what = "character", quiet = TRUE, quote = "\"")
transformed_data <- transformed_data[!grepl("^\\[", transformed_data)]

print(transformed_data)
```


The two exercises below are taken from R for Data Science:

## 3 Describe, in words, what these expressions will match:

- (.)\1\1

This regular expression consists of three parts:
(.): The first part is a capturing group that matches any single character and captures it.
\1: The second part is a backreference to the first capturing group. It matches the same character as captured by the first group.
\1: The third part is another backreference to the first capturing group, matching the same character again.
This regular expression will match three consecutive identical characters.

- (.)(.)\\2\\1

This regular expression consists of four parts:
(.): The first part is a capturing group that matches any single character and captures it.
(.): The second part is another capturing group that matches any single character and captures it.
\\2: This is a backreference to the second capturing group, matching the character captured by the second group.
\\1: This is a backreference to the first capturing group, matching the character captured by the first group.
This regular expression will match two consecutive characters that are the same.

- (..)\1

This regular expression consists of two parts:
(..): The first part is a capturing group that matches any two consecutive characters and captures them.
\1: The second part is a backreference to the first capturing group, matching the same two characters as captured by the first group.
This regular expression will match two consecutive identical pairs of characters.

- (.).\\1.\\1

This regular expression consists of three parts:
(.): The first part is a capturing group that matches any single character and captures it.
\\1: This is a backreference to the first capturing group, matching the character captured by the first group.
\\1: This is another backreference to the first capturing group, matching the same character again.
This regular expression will match a single character surrounded by two identical characters.

- (.)(.)(.).*\\3\\2\\1

This regular expression consists of five parts:
(.): The first part is a capturing group that matches any single character and captures it.
(.): The second part is another capturing group that matches any single character and captures it.
(.): The third part is yet another capturing group that matches any single character and captures it.
.*: This part matches zero or more characters (any character sequence).
\\3\\2\\1: This part is a sequence of backreferences to the capturing groups, matching the characters captured by the third, second, and first groups in that order.
This regular expression will match a sequence of characters where the last three characters are in reverse order compared to the first three characters.


## 4 Construct regular expressions to match words that:

 Start and end with the same character:
 
```{r}
reg1 = "^(.).+\\1"
```

 Contain a repeated pair of letters: 
```{r}
reg2 = "\\b\\w*(\\w{2})\\w*\\1"
```
 

 Contain one letter repeated in at least three places: 
```{r}
reg3 = "^[a-z]*([a-z])\\1[a-z]*"
```
 
 

 
