---
title: "Project 1"
author: "Kossi Akplaka"
date: "`r Sys.Date()`"
output: 
---

```{r echo = FALSE}
knitr::opts_chunk$set(eval = TRUE, message = FALSE, warning = FALSE)
```


## Getting started
### Load packages
Let's load these packages
```{r load-packages, message=FALSE}
library(tidyverse)
library(stringr)
library(knitr)
```

### Import the chess data from github
```{r warning=FALSE}
chess_df = read_csv(file = 'https://raw.githubusercontent.com/Kossi-Akplaka/Data607-data_acquisition_and_management/main/Project%201/chess.txt', col_names = FALSE)
```

Let's have a glimpse of the data
```{r}
head(chess_df,10)
```
## Preprocess the Data
We can drop the first 4 rows of the data and process the rest of the data. The way the data is structured, we can extract:

- Player Number (Useful to find the average), Player’s Name, and the opponents of the players in one line.

- Player’s State, Total Number of Points, Player’s Pre-Rating on the other line

In this part, we can extract those information into 2 different dataset


### Dataset 1: Extract every line starting by a number
```{r}
chess <- matrix(unlist(chess_df), byrow=TRUE)
chess_df1 <- chess[seq(5,length(chess),3)]
head(chess_df1)
```
### Extract data starting by Player's state
```{r}
chess_df2 <- chess[seq(6,length(chess),3)]
head(chess_df2)
```

## Data manipulation
Now, let's extract some useful information

Extract the player's number
```{r}
PairNum <- as.numeric(str_extract(chess_df1, '\\d+'))
```

Extract the player's full name (matching first combination of letters until |, then extract the full name)
```{r}
Name <- str_extract(chess_df1, '[A-z].{1,32}') 
Name <- str_trim(str_extract(Name, '.+\\s{2,}'))
```

Extract the player's state (matching the first 2 letters)
```{r}
State <- str_extract(chess_df2, '[A-Z]{2}') 
```

Extract the player's pre-rating (matching any combination starting from R until -, then taking the first 4 numbers)
```{r}
PreRating <- str_extract(chess_df2, 'R:.{8,}-')
PreRating <- as.numeric(str_extract(PreRating, '\\d{1,4}'))
```

Extract the player's rounds (combination of letter, space, 2 numbers, then taking the numbers)
```{r warning=FALSE}
Rounds <- str_extract_all(chess_df1, '[A-Z]\\s{2,}\\d+')
Rounds <- str_extract_all(Rounds, '\\d+')
```

Extract the total number of points (combination of 1 number, period, a number)
```{r}
TotalNumberofPoints <- as.numeric(str_extract(chess_df1, '\\d+\\.\\d'))
```

Calculate the average Pre chess rating of opponents
```{r}
average_Opponents <- c()
for(i in c(1:length(Rounds))){
  average_Opponents[i] <- round(mean(PreRating[as.numeric(Rounds[[i]])]),0)
}
```

## Creating a new dataframe 

```{r}
chess_dataframe <- data.frame(PairNum,Name,State,TotalNumberofPoints,PreRating,average_Opponents)
kable(head(chess_dataframe,10), row.names = NA)
```
## Load into a csv format
We can load this data frame into a csv format. 
```{r}
write_csv(chess_dataframe, 'chess.csv')
```

