---
title: "Assignment 7"
author: "Kossi Akplaka"
date: "2023-10-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Create data 

I created the data in the notebook and load that to my Github

## Load the data

To load the information from the HTML, XML, and JSON sources into separate R data frames, you can use the following code. We will use the rvest package for HTML, XML and xml2 package for XML, and rjson for JSON.

```{r libraries}
library(rvest) # load html file
library(xml2)  # Load xml file
library(XML)   
library(rjson) # load json file
```


This article "https://www.datacamp.com/tutorial/r-data-import-tutorial" on dataCamp explains how to import HTML, XML, and JSON data in R.

### Load data from HTML into a data frame

Extracting HTML data from github into table

```{r}
file<-read_html("https://github.com/Kossi-Akplaka/Data607-data_acquisition_and_management/blob/main/assignment7/books.html")
table<-html_nodes(file, "table")
table
```
Transform into a dataframe

```{r}
books_html <-html_table(table)[[1]] #converting HTML tables to dataframe
books_html
```

### Load data from XML into a data frame

Getting the data

```{r}
xml_file <- xmlParse(read_xml('https://raw.githubusercontent.com/Kossi-Akplaka/Data607-data_acquisition_and_management/main/assignment7/book.xml'))
xml_file
```
Transform the data into a dataframe

```{r}
books_xml <- xmlToDataFrame(xml_file)
books_xml
```

### Load data from JSON into a data frame

Get the data 

```{r}
json_file <- fromJSON(file = 'https://raw.githubusercontent.com/Kossi-Akplaka/Data607-data_acquisition_and_management/main/assignment7/books.json')
json_file
```
Transform to a dataframe

```{r}
# stack list on top of each other
books_json <- do.call(rbind, lapply(json_file, as.data.frame))
books_json
```

The data frames are identical...
